Concurrent Programming in Harmony

### 2021-12-16 book/paper.tex
Latest

\index{reserve debugging}%
It is time to start stepping through the code that has been executed
before this happened.  This is sometimes known as
\emph{reverse debugging}.  In fact, Harmony allows you to step through
code forwards and backwards.  In this case, we first want to see what
T2 is doing.  You can click on its first (top-left) orange box to
time-travel to that part in the execution.  Now by hitting
$<$return$>$ repeatedly, we can quickly skip through the code.
T2 first calls \texttt{put}(?$q$, 1) and then allocates a new
node initialized with a lock.  Keep stepping until it executes
$\mathit{nq} = q$.  Hit $<$return$>$ once more and inspect the
state of T2 in the lower-right corner.  You can see that variable
\textit{nq} is initialized to~?$q$.  T2 then enters into the
\textbf{while} loop and tries to acquire
$\mathit{nq}$$\rightarrow$$\mathtt{lock}$.  This succeeds, and
next T2 executes \textbf{let} $n = \mathit{nq}$$\rightarrow$$\mathtt{next}$.
Now $n = ?\mathtt{alloc}.\mathit{alloc}[0]$, which is not \texttt{None}.
It then releases $\mathit{nq}$$\rightarrow$$\mathtt{lock}$
(\textit{nq} points to $q$).  It then sets \textit{nq} to $n$, which
is still $?\mathtt{alloc}.\mathit{alloc}[0]$.  Finally, it calls
$\mathtt{acquire}(?\mathit{nq}$$\rightarrow$$\mathtt{lock})$.
But before it can complete that operation, T3 runs next.

T3 chooses "get" and then goes on to invoke $\mathtt{get}(?q)$.
This first successfull acquires $q$$\rightarrow$$\texttt{lock}$.
It then finds out that $q$$\rightarrow$$\texttt{next}$ points to
$\texttt{alloc}.\textit{pool}[0]$.  It sets \textit{node} to
$\texttt{alloc}.\textit{pool}[0]$ as well and sets
$q$$\rightarrow$$\texttt{next}$ to 
$\mathtt{node}$$\rightarrow$$\texttt{next}$.

XXXXXXXX

We can start by single stepping through the
actions of \texttt{putter}(0) by clicking on its first block.
By hitting return repeatedly, we can go through the lines of code that
it is executing.  Doing so, we can see that it allocates
\textit{pool}[0] and uses that as the node to contain value 0 and add
this node to the queue \textit{qtest}.  At the end of its turn,
\texttt{putter}(0) has just finished \texttt{put}(0).  Looking at the
state in the top right, everything looks good.  The first node
(\textit{testq}) points to the allocated node with the value 0 in it,
and its \texttt{next} pointer is \textbf{None}.  The locks on both nodes
are released.  So far so good.

Next \texttt{putter}(1) (thread T2) takes a turn.  It gets to line 10 in
\autoref{fig:queuebroken} where it's trying to acquire
$q$$\rightarrow$\texttt{lock}.  If we look at the bottom right,
we see that $q$ points to \textit{pool}[0], the node that has value 0
in it.  Again, so far so good.

But before \texttt{putter}(1) obtains
the lock on \textit{pool}[0], \texttt{getter}() (thread T4) starts
running.  \texttt{getter}() acquires the lock on \textit{qtest}, which
still points to \textit{pool}[0].  \texttt{getter} then extracts the
value from \textit{pool}[0] and then releases it.
At this point \textit{pool}[0] no longer exists, but then
\texttt{putter}(1) starts running again from where it left off.
It was about to acquire the lock on \textit{pool}[0], but now that
node no longer exists.

>>>>

\index{reserve debugging}%
It is time to start stepping through the code that has been executed
before this happened.  This is sometimes known as
\emph{reverse debugging}.  In fact, Harmony allows you to step through
code forwards and backwards.  In this case, we first want to see what
T2 is doing.  You can click on its first (top-left) orange box to
time-travel to that part in the execution.  Now by hitting
$<$return$>$ repeatedly, we can quickly skip through the code.
T2 first calls \texttt{put}(?$q$, 1) and then allocates a new
node initialized with a lock.  Keep stepping until it executes
$\mathit{nq} = q$.  Hit $<$return$>$ once more and inspect the
state of T2 in the lower-right corner.  You can see that variable
\textit{nq} is initialized to~?$q$.  T2 then enters into the
\textbf{while} loop and tries to acquire
$\mathit{nq}$$\rightarrow$$\mathtt{lock}$.  This succeeds, and
next T2 executes \textbf{let} $n = \mathit{nq}$$\rightarrow$$\mathtt{next}$.
Now $n =\,?\mathtt{alloc}.\mathit{alloc}[0]$, which is not \texttt{None}.
It then releases $\mathit{nq}$$\rightarrow$$\mathtt{lock}$
(\textit{nq} points to $q$).  It then sets \textit{nq} to $n$, which
is still $?\mathtt{alloc}.\mathit{alloc}[0]$.  Finally, it calls
$\mathtt{acquire}(?\mathit{nq}$$\rightarrow$$\mathtt{lock})$.
But before it can complete that operation, T3 runs next.

T3 chooses "get" and then goes on to invoke $\mathtt{get}(?q)$.
This first successfully acquires $q$$\rightarrow$$\texttt{lock}$.
T3 then finds out that $q$$\rightarrow$$\texttt{next}$ points to
$\texttt{alloc}.\textit{pool}[0]$.  T3 sets \textit{node} to
$\texttt{alloc}.\textit{pool}[0]$ as well and sets
$q$$\rightarrow$$\texttt{next}$ to 
$\mathtt{node}$$\rightarrow$$\texttt{next}$.
T3 sets \textit{result} to 
$\mathtt{node}$$\rightarrow$$\texttt{value}$ (which is 1)
and then frees \textit{node}.  This is where the problem is---T2 is
about to acquire the lock in that same node.
